
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

setwd("/EPHOR/Task4.2.3")
getwd()
output.path <-"/EPHOR/Task4.2.3/Rdata/"

library(stabs)
library(glmnet)
library(tictoc)
library(EValue)
library(brms)
library(bayesplot)
library(tidybayes)
library(tidyverse)
library(ggridges)
library(dplyr)
library(metafor)
library(lme4)
library(merTools)

```

```{r file loading, include=FALSE}

#load(file=paste0(output.path,"final.subjects.agnostic4.3d copy.Rdata")) #17:347 are duration variables, 348:659 are dc variables, for penalization model

#load(file=paste0(output.path,"final.subjects.longest.job copy.Rdata")) # file with 3d duration and dc, plus longest held job and longest held years

#load(file=paste0(output.path,"final.subjects.exposome6.Rdata")) #cleaned 4-digit penalization model file

```

```{r data matrix prepartion - load and start from here, include=FALSE}

load(file=paste0(output.path,"final.subjects.agnostic4.3d copy.Rdata")) #17:347 are duration variables, 348:659 are dc variables, for penalization model


all.subjects.isco<-all.subjects.agnostic

is.factor(all.subjects.isco$isco3_011_dc)
table(all.subjects.isco$isco3_011_dc)

for(i in c(348:659)){
  all.subjects.isco[,i]<-as.factor(all.subjects.isco[,i]) 
}
is.factor(all.subjects.isco$isco3_011_dc)

levels(all.subjects.isco$study_name)<-1:20
table(all.subjects.isco$study_name)
table(all.subjects.agnostic$study_name)

all.subjects.isco$A<-as.factor(all.subjects.isco$A)
all.subjects.isco$A[is.na(all.subjects.isco$A)]<-0 #ran a check and all were non-jobs
anyNA(all.subjects.isco$A)
table(all.subjects.isco$A)
all.subjects.isco$sex<-as.factor(all.subjects.isco$sex)


x.all<-all.subjects.isco[,c(2,6:659)] #employment duration + dc variables
x.dur<-all.subjects.isco[,c(2,6:347)] #duration only
x.dc<-all.subjects.isco[,c(2,6:16,348:659)] #duration only

#x.mat<-as.matrix(x.all)
#x.mat<-as.matrix(x.dur)
#x.mat<-as.matrix(x.dc)
y<-as.factor(all.subjects.isco[,4])
```

##Penalization model with dichotomous (dc) variables

```{r Penalization model with dichotomous (dc) variables, include=TRUE}

table(y) #status - 16901 cases; 20965 controls
head(names(x.dc),25) #324 variables, 1:12 are "must have" variables, rest ever employment in 3d isco68 occ

set.seed(1234)

x.mat<-as.matrix(x.dc)

pfac=rep(1,324)
pfac[c(1:12)]=0

dc.model1 <- stabsel(x = x.mat, y = y, fitfun = glmnet.lasso, args.fitfun = list(type = "anticonservative", penalty.factor = pfac, family = "binomial"), cutoff = 0.75, PFER = 10) 

dc.model1$selected



```

```{r Other penalization settings, include=F}

x.dc<-x.dc[,-5] #remost list A jobs
x.mat<-as.matrix(x.dc)

pfac=rep(1,323)
pfac[c(1:11)]=0

dc.model2 <- stabsel(x = x.mat, y = y, fitfun = glmnet.lasso, args.fitfun = list(type = "anticonservative", penalty.factor = pfac, family = "binomial"), cutoff = 0.75, PFER = 10) 

dc.model2$selected

dc.model3 <- stabsel(x = x.mat, y = y, fitfun = glmnet.lasso, args.fitfun = list(type = "anticonservative", penalty.factor = pfac, family = "binomial"), cutoff = 0.75, PFER = 5) 

dc.model3$selected

dc.model4 <- stabsel(x = x.mat, y = y, fitfun = glmnet.lasso, args.fitfun = list(type = "anticonservative", penalty.factor = pfac, family = "binomial"), cutoff = 0.75, PFER = 2) 

dc.model4$selected
```

```{r Logistic regression with selected variables, include=TRUE, message=FALSE}
model1_var<-c("status", names(dc.model1$selected))
model1.selected<-all.subjects.isco[model1_var]
selected.data<-model1.selected

stabs1<-glm(status~., data = selected.data, family = binomial)

AIC(stabs1) 

summary(stabs1)

stabs1.or<-round(exp(cbind(summary(stabs1)$coefficients[36:54,1],confint(stabs1)[36:54,])),3)
stabs1.or
```


```{r Case number for job groups, include=F}
table(all.subjects.isco[all.subjects.isco$isco3_027_dc==1,]$status) #2
table(all.subjects.isco[all.subjects.isco$isco3_072_dc==1,]$status) #83
table(all.subjects.isco[all.subjects.isco$isco3_083_dc==1,]$status) #16
table(all.subjects.isco[all.subjects.isco$isco3_132_dc==1,]$status) #231
table(all.subjects.isco[all.subjects.isco$isco3_193_dc==1,]$status) #69
table(all.subjects.isco[all.subjects.isco$isco3_212_dc==1,]$status) #190
table(all.subjects.isco[all.subjects.isco$isco3_310_dc==1,]$status) #565
table(all.subjects.isco[all.subjects.isco$isco3_331_dc==1,]$status) #624
table(all.subjects.isco[all.subjects.isco$isco3_393_dc==1,]$status) #1198
table(all.subjects.isco[all.subjects.isco$isco3_951_dc==1,]$status) #831
table(all.subjects.isco[all.subjects.isco$isco3_954_dc==1,]$status) #535
table(all.subjects.isco[all.subjects.isco$isco3_999_dc==1,]$status) #1466
table(all.subjects.isco[all.subjects.isco$isco3_AA5_dc==1,]$status) #630

```

```{r E-values for select OR results, include=TRUE}

evalues.OR(1.289,1.118,1.486,rare = TRUE) #for isco-954 carpenters
evalues.OR(1.413,1.238,1.615,rare = TRUE) #for isco-951 bricklayers

```

An E-value for unmeasured confounding is minimum strength of association, on the risk ratio scale, that an unmeasured confounder would need to have with both the treatment and the outcome to fully explain away a specific treatment–outcome association, conditional on the measured covariates.

Example with radon as confounder: suppose radon exposure increase the OR of lung cancer by 2, and suppose "ever carpenters" are two times more likely to be exposed to radon than "never carpenters", than the observed effect on carpenters could be explained away by confounding by radon. This works with one confounder or multiple confounders, as long as their combined effect is similar. 

Thoughts?

```{r prep for Penalization model 5-digits, include=F}

load(file=paste0(output.path,"final.subjects.exposome5 copy.Rdata"))

all.subjects.isco4<-all.subjects[,c(which(colnames(all.subjects)=="subjctid"),101:2706)]
all.subjects.exp<-all.subjects.isco[,c(1:16)]
all.subjects.isco4$subjctid<-sub("^0+([0-9]*)$", "\\1", all.subjects.isco4$subjctid) #get rid of leading zero
all.subjects.isco4<-merge(all.subjects.exp,all.subjects.isco4, by = "subjctid")
all.subjects.isco<-all.subjects.isco4

is.factor(all.subjects.isco$isco_01100_dc)
which(colnames(all.subjects.isco)=="isco_01100_dc") #dc starts at col 1314

for(i in c(1320:2622)){
  all.subjects.isco[,i]<-as.factor(all.subjects.isco[,i]) 
}

is.factor(all.subjects.isco$isco_01100_dc)

levels(all.subjects.isco$study_name)<-1:20
table(all.subjects.isco$study_name)

all.subjects.isco$A<-as.factor(all.subjects.isco$A)
all.subjects.isco$A[is.na(all.subjects.isco$A)]<-0 #ran a check and all were non-jobs
anyNA(all.subjects.isco$A)
table(all.subjects.isco$A)
all.subjects.isco$sex<-as.factor(all.subjects.isco$sex)

# save(all.subjects.isco,file=paste0(output.path,"final.subjects.exposome6.Rdata")) #cleaned 4-digit file

#x<-all.subjects.isco[,c(2,6:2622)] #employment duration + dc variables
x<-all.subjects.isco[,c(which(colnames(all.subjects.isco)=="sex"),
                      which(colnames(all.subjects.isco)=="study_name"),
                      which(colnames(all.subjects.isco)=="packyrs"),
                      which(colnames(all.subjects.isco)=="time_quit"),
                      which(colnames(all.subjects.isco)=="agegroup"),
                      which(colnames(all.subjects.isco)=="A"),
                      which(colnames(all.subjects.isco)=="silica.cumexposure.lag0"),
                      which(colnames(all.subjects.isco)=="asbestos.cumexposure.lag0"),
                      which(colnames(all.subjects.isco)=="dme.cumexposure.lag0"),
                      which(colnames(all.subjects.isco)=="pah.cumexposure.lag0"),
                      which(colnames(all.subjects.isco)=="nickel.cumexposure.lag0"),
                      which(colnames(all.subjects.isco)=="chromium.cumexposure.lag0"),
  1320:2622)] #dichotomous only

x.mat<-as.matrix(x)
y<-as.factor(all.subjects.isco[,4])
```


```{r 5-digit penalization model, include=FALSE}
table(y) #status - 16901 cases; 20965 controls
head(names(x),25) #1315 variables, 1:12 are "must have" variables, rest ever employment in 5d isco68 occ

set.seed(1234)

pfac=rep(1,1315)
pfac[c(1:12)]=0

dc.model1_4d <- stabsel(x = x.mat, y = y, fitfun = glmnet.lasso, args.fitfun = list(type = "anticonservative", penalty.factor = pfac, family = "binomial"), cutoff = 0.75, PFER = 10) 

dc.model1_4d$selected
```

```{r Logistic regression with selected variables - 4d, include=F, message=FALSE}
model1.4_var<-c("status", names(dc.model1_4d$selected))
model1.4.selected<-all.subjects.isco[model1.4_var]
selected.data<-model1.4.selected

stabs1.4<-glm(status~., data = selected.data, family = binomial)

AIC(stabs1.4) 

summary(stabs1.4)

stabs1.4or<-round(exp(cbind(summary(stabs1.4)$coefficients[36:57,1],confint(stabs1.4)[36:57,])),3)
stabs1.4or
```

```{r Case number for job groups - 5d, include=F}
table(all.subjects.isco[all.subjects.isco$isco_02410_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_03550_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_07210_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_08310_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_13215_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_13290_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_19390_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_21210_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_31010_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_39310_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_43130_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_83290_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_95120_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_95160_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_95410_dc==1,]$status) #
table(all.subjects.isco[all.subjects.isco$isco_97155_dc==1,]$status) #
```




